---
title: "STL Lead Inequity"
author: "Wylie Hampson"
date: "8/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(skimr)
```

## Read in the stl_blood_lead.csv as a data frame

read_delim() is for csv files that have different delimiters

```{r}
stl_lead <- read_delim(here::here("data", "stl_blood_lead.csv"), delim = ";")

stl_lead <- stl_lead %>% clean_names()

```

Create new data frame that copies stl_lead, but with one extra column called prop_white

```{r}
stl_lead_prop <- stl_lead %>% 
  dplyr::mutate(prop_white = (white / total_pop) * 100) %>% 
  dplyr::mutate(pct_elevated = str_replace(pct_elevated, pattern = ",", replace = ".")) %>% 
  dplyr::mutate(pct_elevated = as.numeric(pct_elevated))

```

## Create a scatterplot of the percentage of children in each census tract with elevated blood levels versus the percent of each census tract tract identifying as white.

```{r, fig.width = 5, fig.height = 5}
stl_lead_plot <- ggplot(data = stl_lead_prop, aes(x = prop_white, y = pct_elevated)) +
  geom_point(alpha = 0.5,
             color = "purple") +
  theme_light() +
  labs(x = "Proportion White",
       y = "Percent of Elevated Lead Levels",
       title = "Percent of Elevated Lead Blood Levels vs. Percent White")

stl_lead_plot
```

## Save the plot to the figs file

```{r}
ggsave(here::here("figs", "stl_lead_plot.jpg"), plot = stl_lead_plot, width = 6, height = 5)
```

The above graph shows that as the proportion of communities becomes increasingly white, the percentage of elevated lead levels in children becomes lower.

## Make a histogram

```{r, fig.width = 7, fig.height = 7}
stl_lead_histogram <- ggplot(data = stl_lead_prop, aes(x = pct_elevated)) +
  geom_histogram(aes(fill = "yellow"),
                 alpha = 0.5,
                 color = "brown") +
  theme_dark() +
  labs(x = "Percentage of Elevated Blood Lead Level",
       y = "Count",
       title = "Kids with Elevated Blood Lead Levels by Tract")

stl_lead_histogram
```

## Save the histogram to the figs folder

```{r}
ggsave(here::here("figs", "stl_lead_histrogram.jpg"), plot = stl_lead_histogram,
       width = 6, height = 5)
```

